Multiple and Logistic Regression

### 2018-11-28 DATA606/Multiple and Logistic Regression.Rmd
Work on week 8 assignment

(a) Write the equation of the regression line.

(b) Interpret the slope in this context, and calculate the predicted birth weight of first borns and others.

(c) Is there a statistically significant relationship between the average birth weight and parity?

@@

The summary table below shows the results of a linear regression model for predicting the average number of days absent based on ethnic background (eth: 0 - aboriginal, 1 - not aboriginal), sex (sex: 0 - female, 1 - male), and learner status (lrn: 0 - average learner, 1 - slow learner).1

@@

(a) Write the equation of the regression line.

(b) Interpret each one of the slopes in this context.

@@

@@

@@

Which, if any, variable should be removed from the model first?

@@

(a) Each column of the table above represents a different shuttle mission. Examine these data and describe what you observe with respect to the relationship between temperatures and damaged O-rings.

@@

@@

(d) Based on the model, do you think concerns regarding O-rings are justified? Explain

@@

$log(\frac{\hat{P}}{1- \hat{P}}) = 11.6630 − 0.2162 \times \text{Temperature}$

where $\hat{p}$ is the model-estimated probability that an O-ring will become damaged. Use the model to calculate the probability that an O-ring will become damaged at each of the following ambient temperatures: 51, 53, and 55 degrees Fahrenheit. The model-estimated probabilities for several additional ambient temperatures are provided below, where subscripts indicate the temperature:
$\hat{p}_\text{57}$ = 0.341 $\hat{p}_\text{59}$ = 0.251 $\hat{p}_\text{61}$ = 0.179 $\hat{p}_\text{63}$ = 0.124
$\hat{p}_\text{65}$ = 0.084 $\hat{p}_\text{67}$ = 0.056 $\hat{p}_\text{69}$ = 0.037 $\hat{p}_\text{71}$ = 0.024

@@

@@

>>>>

(a) Write the equation of the regression line. $\widehat{weight}=120.07 - 1.93 \times parity$

(b) Interpret the slope in this context, and calculate the predicted birth weight of first borns and others. **All else being equal, the average birth weight is 120 ounces.  If the child is not the first born the birth weight decreases by two ounces, to 118 ounces.**

(c) Is there a statistically significant relationship between the average birth weight and parity? **No because the P value is > 0.05**

@@

The summary table below shows the results of a linear regression model for predicting the average number of days absent based on ethnic background (eth: 0 - aboriginal, 1 - not aboriginal), sex (sex: 0 - female, 1 - male), and learner status (lrn: 0 - average learner, 1 - slow learner).

@@

(a) Write the equation of the regression line. $\widehat{day} = 18.93 -9.11eth + 3.1sex + 2.15lrn$

(b) Interpret each one of the slopes in this context. **On average a student misses about 19 days of school.  If they are aboriginal they miss an additional 9 days.  If they are female they miss 3 more days.  If they are not a slow learner they miss 2 additional days.**

@@

$\widehat{day} = 18.93 -9.11(0) + 3.1(1) + 2.15(1) = 24.18$

$e_i = day_1 - \widehat{day}_1 = 2 - 24.18 = -22.18$

@@

$R^2 = 1 - \frac{Var(e_1)}{Var(y_1)} = 1 - \frac{240.57}{264.17} = 0.0893364$

$R^2_\text{adj} = 1 - \frac{Var(e_1)}{Var(y_1)} \times \frac{(n-1)}{(n-k-1)} = 1- \frac{240.57}{264.17} \times \frac{(146-1)}{(146 - 3 - 1)} = 1 - 0.9106636 \times 1.021127 = 0.07009681$

@@

Which, if any, variable should be removed from the model first? **lrn because the adjusted** $R^2$ **increases from 0.0701 to 0.0723**

@@

(a) Each column of the table above represents a different shuttle mission. Examine these data and describe what you observe with respect to the relationship between temperatures and damaged O-rings. **At 53 to 57 degrees Fahrenheit, the ratio of damaged to undamaged O-rings flip.**

@@

**The model is the probability of failure given the ambient air temperature at launch time.**

@@

$log(\frac{p}{1-p}) = 11.6630 - 0.2162 \times \text{Temperature}$

(d) Based on the model, do you think concerns regarding O-rings are justified? Explain.

**Yes.  Temperature is a statistically significant variable.  To see this effect consider the following:  What is the probability of failure when the launch temperature is 81 degrees?  What is the probability of failure at 53?**

$P = \frac{e^L}{1 + e^L}$ **where**
$L = 11.6630 - 0.2162 \times \text{Temperature}$

```{r}
probability_of_failure <- function(temperature){
  l <- 11.6630 - 0.2162 * temperature
  round(exp(l) / (1 + exp(l)), 3)
}

probability_of_failure(81)
probability_of_failure(53)
```

**The probability of failure increases from 0 to 0.6 within the observed extreme temperatures.**

@@

$log(\frac{\hat{p}}{1- \hat{p}}) = 11.6630 − 0.2162 \times \text{Temperature}$

where $\hat{p}$ is the model-estimated probability that an O-ring will become damaged. Use the model to calculate the probability that an O-ring will become damaged at each of the following ambient temperatures: 51, 53, and 55 degrees Fahrenheit.

* $\hat{p}_\text{51}$ = `r probability_of_failure(51)`
* $\hat{p}_\text{53}$ = `r probability_of_failure(53)`
* $\hat{p}_\text{55}$ = `r probability_of_failure(55)`

The model-estimated probabilities for several additional ambient temperatures are provided below, where subscripts indicate the temperature:

* $\hat{p}_\text{57}$ = 0.341 
* $\hat{p}_\text{59}$ = 0.251 
* $\hat{p}_\text{61}$ = 0.179 
* $\hat{p}_\text{63}$ = 0.124
* $\hat{p}_\text{65}$ = 0.084 
* $\hat{p}_\text{67}$ = 0.056 
* $\hat{p}_\text{69}$ = 0.037 
* $\hat{p}_\text{71}$ = 0.024

@@

```{r}
library(ggplot2)
temp <- seq( 51, 71, 2)
p <- probability_of_failure(temp)
df <- data.frame(temperature = temp, probability_of_failure = p)

spline_int <- as.data.frame(spline(df$temp, df$probability_of_failure))

ggplot(df) + 
  geom_point(aes(x = temp, y = probability_of_failure), size = 3) +
  geom_line(data = spline_int, aes(x = x, y = y)) +
  labs(x = "Temperature (degrees Fahrenheit)", y = "Probability of Failure")
```

@@

**This model may underestimate the risk at the higher temperatures.  From the observational data we see 2 of 3 cases at 70 degrees had a failure.  The model would predict `r probability_of_failure(70)`.  Since we are dealing with human lives an abundance of caution should be exercised.**

**There are two assumptions required to accept the model's validity: Each outcome is independent of the other outcomes and each predictor is linearly related if all other predictors are held constant.  The outcomes might not be independent as different manufacturing processes might be employed in response to a previous shuttle launch.  It also doesn't appear that the risk of failure and the temperature are linearly related.**
